\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb}
\usepackage{enumitem}
\usepackage{tikz}
\usepackage{graphicx}
\usepackage{float}
\usepackage{placeins}
\usepackage{booktabs}

\title{HW \#5}
\author{Theo Tarr}
\date{March 5, 2026}

\begin{document}
\maketitle

\section*{1. Reading-Related Question}

\textbf{(a)} The \textit{primitive soup} hypothesis (Oparin–Haldane) proposes that, in a primitive oxygen-free atmosphere, UV light and lightning discharges synthesized a wide range of organic compounds — amino acids, sugars, purines, pyrimidines — which accumulated in the sea until it reached the consistency of a ``hot dilute soup'' (Ch.~3). A key weakness is that the reactions lack the necessary \textit{specificity}. Miller-type experiments produce organic compounds, but forming biological polymers requires very particular chemical linkages: proteins need peptide bonds, and RNA requires nucleotides to be joined in exactly the right way. If any linkages are of the wrong kind, replication is blocked. The soup offers no mechanism to enforce this specificity, and long-chain fatty acids needed for membranes were absent entirely.

The \textit{primitive pizza} hypothesis (Wächtershäuser) proposes instead that life began on charged mineral surfaces — specifically positively charged iron pyrites (fool's gold) — to which organic ions became bonded (Ch.~3). Because the molecules are held in a fixed orientation and constrained to move only across the surface in two dimensions rather than freely in solution, both the speed and the specificity of chemical reactions are greatly increased. Surface bonding also raises the local concentration of interacting molecules, solving the dilution problem. This directly addresses the weakness of the soup: the surface enforces the correct molecular orientations required to produce properly-linked polymers.

\textbf{(b)} As described in Chapter~4, RNA replication does not benefit from the proofreading mechanisms available to DNA, because the new and old strands separate immediately upon synthesis, making it impossible to check whether base pairing was correct. As a result, the error rate in RNA replication (with replicase, but without proofreading) is in the range $1/1000$ to $1/10{,}000$ per nucleotide. Without replicase enzymes, the error rate is considerably higher still — replication would be so inaccurate that almost no faithful copies could be produced, severely limiting the length of any molecule that could be reliably replicated (the error threshold, Ch.~3). This is consistent with expectations: enzymes improve specificity, but without proofreading the error rate remains too high for large genomes.

\textbf{(c)} As Chapter~4 explains, a key difference between RNA and DNA is that RNA is normally single-stranded, whereas DNA exists as a double helix. This single-stranded nature has two critical consequences for a human with an RNA genome. First, proofreading is impossible: there is no complementary strand against which to check for mismatches, so errors introduced during replication cannot be corrected. With an error rate of $1/1000$ to $1/10{,}000$ per nucleotide and a human genome of roughly 3 billion bases, every single round of DNA replication would introduce hundreds of thousands to millions of mutations — far beyond any viable error threshold. Second, RNA is chemically less stable than DNA: the extra $2'$-OH group on the ribose backbone makes RNA susceptible to spontaneous hydrolysis. The accumulated effect of rampant mutation and rapid chemical degradation would be error catastrophe: the genetic information would be destroyed faster than it could be replicated, and the organism could not survive.

\section*{2. $^{12}$C--$^{13}$C Fractionation}

\textbf{(a)} Using $m_\text{proton} = m_\text{neutron} = 1.67 \times 10^{-27}$ kg:
\[
m_{^{12}\text{C}} = 12 \times 1.67\times10^{-27}\ \text{kg} = 2.004\times10^{-26}\ \text{kg}
\]
\[
m_{^{13}\text{C}} = 13 \times 1.67\times10^{-27}\ \text{kg} = 2.171\times10^{-26}\ \text{kg}
\]

\textbf{(b)} For a particle in a box of width $L = 0.2\ \text{nm} = 2\times10^{-10}\ \text{m}$:
\[
E_n = \frac{n^2 h^2}{8mL^2}
\]

For $n=1$:
\[
E_1 = \frac{h^2}{8mL^2} = \frac{(6.626\times10^{-34})^2}{8 \times m \times (2\times10^{-10})^2}
= \frac{4.390\times10^{-67}}{8 \times m \times 4\times10^{-20}}
\]
\[
E_1(^{12}\text{C}) = \frac{4.390\times10^{-67}}{6.413\times10^{-45}} = 6.85\times10^{-23}\ \text{J}\ \approx 0.427\ \text{meV}
\]
\[
E_1(^{13}\text{C}) = \frac{4.390\times10^{-67}}{6.947\times10^{-45}} = 6.32\times10^{-23}\ \text{J}\ \approx 0.394\ \text{meV}
\]

The $n=1$ wavefunction is a half-sine with no internal nodes; both isotopes have the same wave shape. Because $^{12}$C is lighter, it has a higher ground-state (zero-point) energy.

\begin{center}
\begin{tikzpicture}[scale=1.1]
  % Box walls
  \draw[thick] (0,0) -- (0,2.5);
  \draw[thick] (4,0) -- (4,2.5);
  \draw[thick] (0,0) -- (4,0);

  % 12C energy level (higher)
  \draw[dashed, gray] (0,1.8) -- (4,1.8);
  \node[right] at (4.05,1.8) {\footnotesize $E_1(^{12}\text{C})$};
  % 12C wave
  \draw[blue, thick, domain=0:4, samples=60]
    plot (\x, {0.5*sin(deg(3.14159*\x/4)) + 1.8});

  % 13C energy level (lower)
  \draw[dashed, gray] (0,1.2) -- (4,1.2);
  \node[right] at (4.05,1.2) {\footnotesize $E_1(^{13}\text{C})$};
  % 13C wave
  \draw[red, thick, domain=0:4, samples=60]
    plot (\x, {0.4*sin(deg(3.14159*\x/4)) + 1.2});

  \node[blue] at (2, 2.45) {\footnotesize $^{12}$C};
  \node[red] at (2, 0.75) {\footnotesize $^{13}$C};

  \node[below] at (2, -0.15) {\footnotesize $L = 0.2\ \text{nm}$};
\end{tikzpicture}
\end{center}

\textbf{(c)} Let $E_\text{ts}$ be the (same) transition-state energy for both isotopes. The activation energy for each is:
\[
E_a = E_\text{ts} - E_1
\]
Since rate $\propto e^{-E_a/k_BT}$, the ratio is:
\[
\frac{\text{rate}(^{12}\text{C})}{\text{rate}(^{13}\text{C})}
= \frac{e^{-(E_\text{ts}-E_1^{12})/k_BT}}{e^{-(E_\text{ts}-E_1^{13})/k_BT}}
= e^{(E_1^{12}-E_1^{13})/k_BT}
\]
\[
\Delta E = E_1(^{12}\text{C}) - E_1(^{13}\text{C}) = 6.85\times10^{-23} - 6.32\times10^{-23} = 5.3\times10^{-24}\ \text{J}
\]
\[
k_BT = (1.381\times10^{-23})(300) = 4.14\times10^{-21}\ \text{J}
\]
\[
\frac{\Delta E}{k_BT} = \frac{5.3\times10^{-24}}{4.14\times10^{-21}} = 1.28\times10^{-3}
\]
\[
\frac{\text{rate}(^{12}\text{C})}{\text{rate}(^{13}\text{C})} = e^{1.28\times10^{-3}} \approx \boxed{1.00128}
\]
$^{12}$C reacts about $0.13\%$ faster than $^{13}$C, consistent with the observed kinetic isotope fractionation in photosynthesis.

\textbf{(d)} At higher temperature, $k_BT$ increases, so $\Delta E/k_BT$ decreases and the rate ratio approaches 1. The fractionation effect is reduced: both isotopes react at nearly the same rate, and the $^{12}$C/$^{13}$C discrimination by plants weakens.

\textbf{(e)} For $^{13}$C to react faster, its activation energy must be smaller: $E_a(^{13}\text{C}) < E_a(^{12}\text{C})$. Since $E_1(^{12}\text{C}) > E_1(^{13}\text{C})$, this requires the \textit{transition state} to also have quantized zero-point energy, with $E_\text{ts}(^{12}\text{C})$ exceeding $E_\text{ts}(^{13}\text{C})$ by \textit{more} than the difference in their ground-state energies. This occurs when the bond to the carbon atom is \textit{tighter} (stiffer) in the transition state than in the reactant, giving $^{12}$C a larger zero-point energy penalty at the transition state. This is the origin of inverse kinetic isotope effects.

\section*{3. First Cells}

\textbf{(a)} The RNA strand inside the vesicle replicates, consuming free nucleotides and incorporating them into large polymers. Since the membrane is impermeable to polymers, the RNA chains cannot leave. This lowers the free nucleotide concentration inside the vesicle relative to outside, creating a concentration gradient. Single nucleotides then diffuse down this gradient, flowing inward faster than they exit, until equilibrium is approached.

\textbf{(b)} Water moves by osmosis from regions of low solute concentration (high water activity) to regions of high solute concentration (low water activity). A higher total concentration of RNA polymers inside the vesicle means the interior has lower water activity than the exterior. Water therefore flows into the vesicle until the osmotic pressure is balanced by the membrane tension.

\textbf{(c)} Vesicles with higher RNA concentrations have a greater osmotic pressure drawing water inward, so they swell and grow. To accommodate this growth, they also absorb fatty acid molecules from solution, pulling them away from vesicles with lower RNA concentrations. The low-RNA vesicles therefore lose membrane material and shrink. This creates competition among vesicles: those that replicate RNA more effectively grow at the expense of less productive ones. This is a form of natural selection at the level of primitive cells — a Darwinian process in which "fitter" (more productive) vesicles propagate, directly linking RNA replication fidelity to vesicle survival and growth.

\textbf{(d)} The first cells had no helicase enzymes or other active machinery to separate the double-stranded RNA. Strand separation could instead be achieved by \textit{thermal cycling}: hydrothermal environments cycle between high and low temperatures, and at elevated temperatures the two strands would denature (melt apart) as hydrogen bonds break. Upon cooling, each single strand could then act as a template for a new complementary strand. This passive, thermally-driven mechanism requires no molecular machinery and is consistent with a prebiotic hydrothermal origin scenario.

\section*{4. Handedness}

\textbf{(a)} Using the Boltzmann distribution with $\Delta E = 10^{-17}\ \text{eV}$ (L favored):
\[
\frac{P(L)}{P(D)} = e^{\Delta E / k_BT}
\]
\[
k_BT = (8.617\times10^{-5}\ \text{eV/K})(295\ \text{K}) = 0.02542\ \text{eV}
\]
\[
\frac{\Delta E}{k_BT} = \frac{10^{-17}}{0.02542} = 3.93\times10^{-16}
\]
\[
\frac{P(L)}{P(D)} = e^{3.93\times10^{-16}} \approx 1 + 3.93\times10^{-16} \approx \boxed{1.000\,000\,000\,000\,000\,4}
\]
The two forms are present in essentially equal probability at room temperature — the L-form is favored by only about 1 part in $10^{16}$.

\textbf{(b)} We want $P(D)/P(L) = 0.1$, i.e., $P(L)/P(D) = 10$:
\[
e^{\Delta E / k_BT} = 10
\qquad\Rightarrow\qquad
k_BT = \frac{\Delta E}{\ln 10} = \frac{10^{-17}\ \text{eV}}{2.303} = 4.34\times10^{-18}\ \text{eV}
\]
\[
T = \frac{4.34\times10^{-18}\ \text{eV}}{8.617\times10^{-5}\ \text{eV/K}} \approx \boxed{5.04\times10^{-14}\ \text{K}}
\]
This is an unphysically low temperature, far below any achievable in nature.

\textbf{(c)} No. The relative probability difference from the weak force at room temperature is only $\sim 10^{-16}$, which is far too small to drive the near-complete homochirality ($\sim$100\% L-amino acids) observed in biology. The weak force effect alone could not account for this bias.

\textbf{(d)} An alternative explanation for homochirality is \textit{amplification through autocatalysis}: a small initial chiral imbalance (which could be random, or seeded by circularly polarized UV light from a neutron star that preferentially destroys one enantiomer) is amplified to homochirality through asymmetric autocatalysis, where one chiral form catalyzes its own production while suppressing the other.

\end{document}
